---
title: "HW 01 - Earthquakes"
author: "RUBRIC"
date: "2/5/2023"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
library(openintro)
```

## Exercises

### Exercise 1

MAX 0.5 points
Correct number of rows/earthquakes 0.5 points

There are 123 earthquakes in the dataset.

### Exercise 2

MAX 0.5 points
Correct number of variables

There are 7 variables for each earthquake.

### Exercise 3

Max 0.5 points
Correct countries identified 0.5 points

Turkey, Iran and China

### Exercise 4

MAX 2.5 points
Correct starting code copied 0.5 points
Add pipe operator 0.5 points
Add arrange function 0.5 points
Add richter (with or without desc()) 0.5 points
Include (reasonable) narrative description of results 0.5 points

The code chunk below shows that the most common intensities for major earthquakes are 7.5 and 7.3, but range between 6.8 and 7.5.

```{r earthquakes-top10, eval=TRUE}
earthquakes %>% 
  count(richter, sort = TRUE) %>% 
  slice_max(n, n = 5) %>% 
  arrange(desc(richter))
```

OR

```{r earthquakes-top10, eval=TRUE}
earthquakes %>% 
  count(richter, sort = TRUE) %>% 
  slice_max(n, n = 5) %>% 
  arrange(richter)
```

### Exercise 5

MAX 1 points
0.5 "no relationship"
0.5 "no relationship between magnitude and year" (or similar, but must include variable names)

No relationship. 

Therefore, there is no relationship between earthquake magnitude and year of occurence during the 20th century. 

### Exercise 6

MAX 0.5 points

0.5 No, because the timing of large earthquakes does not seem to correspond to the timing of high death tolls.

### Exercise 7

MAX 0.5 points

0.5 More confident, because this plot shows **more clearly** that high magnitude earthquakes do not always cause large death tolls.