---
title: "HW 02 - Water and Sanitation - Answer Key"
author: "RUBRIC"
date: "Spring 2023"
output: html_document
---

## Load packages and data

```{r load-packages-data, message = FALSE}
library(tidyverse)
wat_san <- read_csv("data/water-and-sanitation.csv")
```

## Exercises

**MAX GRADE: 10 points**

### Exercise 1

MAX 1.5 points
0.5 for nrow
0.5 for ncol
0.5 for in-line code

The are `r nrow(wat_san)` observations and `r ncol(wat_san)` variables in our dataset.

### Exercise 2

MAX 0.5 points
0.5 for intepretation

Each row in the dataset is water and sanitation access stats for one year for a given country.

### Exercise 3

MAX 1.5 points
0.5 for correct code
0.5 for reasonable bin width (needs to be wide >> 1 to see)
0.5 for interpretation

"The distribution has moved closer to zero since 2000, meaning more countries now have 0% of the population with no sanitation access"
OR
"... meaning sanitation access has improved"
OR
"The number of countries where no-sanitation is equal to 0% of the population has increased"

```{r no-sanitation-years, echo = F}
ggplot(data = wat_san, mapping = aes(x = sanitation_none)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~Year)
```

### Exercise 4

MAX 1.5 point
0.5 for correct pipeline
0.5 for slice
0.5 for interpretation

"There appears to be no relationship between zero saniation and zero clean water access"

```{r bottom-5-2000}
wat_san %>% 
  filter(Year == 2000) %>% 
  arrange(desc(sanitation_none)) %>% 
  slice(1:5)
```

### Exercise 5

Max 1 point
0.5 for scatter plot 
0.5 for interpretation

"Yes, this also shows there is no clear relationship, lots of scatter, not a straight line"

```{r all-2000-plot}
wat_san %>% 
  filter(Year == 2000) %>% 
  ggplot(aes(x = water_none, y = sanitation_none)) +
  geom_point()
```

### Exercise 6

Max 1 point
0.5 for correct code (slice optional)
0.5 for interpretation

Yes, several changes:
Added in 2020: Chad, South Sudan, Benin, Namibia
Removed in 2020: India, Ethiopia, Eritrea, Cambodia

```{r bottom-5-2020}
wat_san %>% 
  filter(Year == 2020) %>% 
  arrange(desc(sanitation_none)) %>% 
  slice(1:5)
```

### Exercise 7

Max 1.5 points
0.5 for copying code chunk
0.5 for completing code chunk
0.5 for interpretation

"Overall, water and sanitation access has improved (no access declined), however rates of sanitation_none are still higher than water_none"

```{r progress-summary}
  wat_san %>% 
  filter(Year == 2000 & Entity != "World" | Year == 2020 & Entity != "World") %>% 
  # this line uses a combination of logical operators (== (equal to), & (and), != (is not equal to), | (or))
  # this combination filters for the two years provided, while excluding any data associated with the "World" Entity
  group_by(Year) %>% 
  # this line groups all the data by Year, so that any summary statistics specified below are computed for each group
  summarize(water_none_min = min(water_none, na.rm = T),
            water_none_mean = mean(water_none, na.rm = T),
            water_none_median = median(water_none, na.rm = T),
            water_none_stdev = sd(water_none, na.rm = T),
            water_none_iqr = IQR(water_none, na.rm = T),
            water_none_max = max(water_none, na.rm = T),
            
            sanitation_none_min = min(sanitation_none, na.rm = T),
            sanitation_none_mean = mean(sanitation_none, na.rm = T),
            sanitation_none_median = median(sanitation_none, na.rm = T),
            sanitation_none_stdev = sd(sanitation_none, na.rm = T),
            sanitation_none_iqr = IQR(sanitation_none, na.rm = T),
            sanitation_none_max = max(sanitation_none, na.rm = T)
            )
```

### Exercise 8

Max 1.5 points
0.5 for adding countries to fiter
0.5 for completing aes mappings
0.5 for interpretation

Sanitation_none declined rapidly in India and Cambodia between 2000 and 2020.
Sanitation_none declined slowly in Eritrea and Niger, so this is why Niger stayed in the top five with high Sanitation_none access.

```{r country-global-changes}
wat_san %>% 
  filter(Entity == "World" | Entity == "India" | Entity == "Cambodia" | Entity == "Niger" | Entity == "Eritrea") %>% 
  ggplot(aes(x = Year, y = sanitation_none, color = Entity)) +
  geom_line()
```
