---
title: "Lab 01 - Hello R"
author: "RUBRIC"
date: "2/5/2023"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```

## Exercises

### Exercise 1

MAX 1 point
1846 rows AND 3 variables 0.5 points

dataset, x, y 0.5 points

Any method to output the rows, variables, and variable names is acceptable. 

### Exercise 2

MAX 1 point
Removing text 0.5 points
Leaving the code as-is 0.5 points

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

MAX 3 points
Removing provided text 0.5 points
Adding narrative to describe analysis steps 0.5 points
Correct dataset filtering code 0.5 points 
Correct plotting code 0.5 points 
Correct correlation code 0.5 points
Changing the dataset name to `star` throughout code 0.5 points

First we filter the dinosaurus_dozen dataset for the `star` dataset, then plot using ggplot2

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()
```


```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

### Exercise 4

MAX 2.5 points
Removing provided text 0.5 points
Adding narrative to describe analysis steps 0.5 points
Correct dataset filtering code 0.5 points 
Correct plotting code 0.5 points 
Correct correlation code 0.5 points
Changing the dataset name to `circle` throughout code 0.5 points

First we filter the dinosaurus_dozen dataset for the `circle` dataset, then plot using ggplot2

```{r}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```


```{r}
circle_data %>%
  summarize(r = cor(x, y))
```


### Exercise 5

MAX 2 points
Removing provided text 0.5 points
Adding narrative to describe analysis steps 0.5 points
Copy both code chunks correctly 1 point 
NOTE: should be `eval = T` in first line, but OK if not changed


Let's plot all datasets at once. In order to do this we will make use of faceting.

```{r all-viz, eval=TRUE, fig.fullwidth=TRUE}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point()+
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")
```

And we can use the `group_by` function to generate all the summary correlation coefficients.

```{r all-r, eval=TRUE}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)
```

