---
title: "Lab 02 - Plastic waste"
author: "RUBRIC"
date: "Spring 2023"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read_csv("data/plastic-waste.csv")
```

## Exercises

**MAX GRADE: 10 points**

### Exercise 1

MAX 1.5 point
0.5 for correct dataset, aesethetic and geom
0.5 for correct use of facet_wrap or facet_grid
0.5 for reasonable interpretation 

E.g., reference to outliers (North america), more values close to zero (Africa and/or Oceania), overall similarity among continents  

```{r plastic-waste-continent}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap)) +
  geom_histogram() +
  facet_wrap(~continent)
```

### Exercise 2

MAX 1
0.5 for copying code
0.5 for lowering the alpha value below 0.7

```{r plastic-waste-density}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.2)
```

### Exercise 3

MAX 1 point
0.5 for mapping
0.5 for setting

We want color and fill aesthetics to vary BY continent
We want alpha to be applied (set) for all continents

### Exercise 4

MAX 1.5 points
0.5 for adapting the code for violin plots
0.5 for distribution comment
0.5 for summary stat identification (any of below)

Violin plots (as with density plots) show the distribution of the data, rather than just the summary statistics shown in a box plot.
Box plots show the median AND/OR 1st and 3rd quartile range AND/OR outliers.

```{r plastic-waste-violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 5

MAX 1.5 points
0.5 correct dataset and x/y mapping
0.5 correct gom
0.5 reasonable interpretation

"there is a general positive relatioship between plastic waste per capita and managed waste per capita, but the relationship is messy"

```{r plastic-waste-mismanaged}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

### Exercise 6

Max 1 points
0.5 add color mapping to continent
0.5 reasonable interpretation

"there is no clear by-continent relationship"
"there may be slightly different relationships for each continent, but not easy to see"
"asia and africa seem to have steeper relationships (more unmanaged waste per unit waste) than europe"

```{r plastic-waste-mismanaged-continent}
plastic_waste %>% 
  ggplot(aes(x = plastic_waste_per_cap, 
             y = mismanaged_plastic_waste_per_cap,
             color = continent)) +
  geom_point()
```

### Exercise 7

Max 1.5 points
0.5 for first chunk (correct x/y mapping)
0.5 for second chunk (correct x/y mapping)
0.5 for interpretation

"there seems to be a stronger association with coastal population than with total population"

```{r plastic-waste-population-total}
plastic_waste %>% 
  ggplot(aes(x = total_pop, 
             y = plastic_waste_per_cap,
             color = continent)) +
  geom_point()
```

```{r plastic-waste-population-coastal}
plastic_waste %>% 
  ggplot(aes(x = coastal_pop, 
             y = plastic_waste_per_cap,
             color = continent)) +
  geom_point()
```

### Exercise 8

MAX 1 point
0.5 for copying/recreating the plot accurately
0.5 for reasonable interpreation

"there seems to be a positive relationship between plastic waste per capita and the proprtion of a country's popluation that lives near the coast"

```{r recreate-viz}
plastic_waste %>% 
  mutate(coastal_pop_prop = coastal_pop / total_pop) %>%
  filter(plastic_waste_per_cap < 3) %>%
  ggplot(aes(x = coastal_pop_prop, y = plastic_waste_per_cap, color = continent)) + 
    geom_point() +
    geom_smooth(color = "black") +
    scale_color_viridis_d() +
    labs(x = "Coastal population proportion (Coastal / total population)", 
         y = "Plastic waste per capita ", 
         color = "Continent",
         title = "Plastic waste vs. coastal population proportion",
         subtitle = "by continent") +
    theme_minimal()
```
